---
title: "Elixir Belgium workshop feedback"
subtitle: "January 2018 - March 2018"
author: "Paula Andrea Martinez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tufte)
library(tidyverse)
library(scales)
library(lubridate)
```

```{r echo=FALSE, message=FALSE}
# read in overall data
overall <- read_tsv("data/overall.tsv")
total_attendees <- sum(overall$Attendees)
total_women <- sum(overall$Attendees_Female)
total_men <- sum(overall$Attendees_Male)
total_applicants <- sum(overall$Applicants)
```

# Elixir Belgium workshops

## Introduction

Elixir Belgium Training has the aim to provide tools for data analysis to life 
scientist from all career stages. In this report we would like to share
the feedback received from attendees.

## Summary of attendees

From January 2018 to March 2018 we had a total of **`r nrow(overall)`** 
workshops at different locations around Belgium. 
On these workshops we received **`r total_attendees`** attendees, from 
`r total_applicants` applicants. The gender distribution was 
**`r percent(total_women / total_attendees)` **
for women and **`r percent(total_men / total_attendees)` ** for men.

Number of workshops by location
```{r echo = FALSE}
knitr::kable(
  table(overall$Location), col.names = c("Location", "Number of workshops")
)
```

# Feedback

```{r echo=FALSE, message=FALSE}
# Read in data
feedbackall <- read_csv(file = "data/allworkshops_uptomarch.csv", 
                   na = c(""," ","NA"))
# Clean names
names(feedbackall) <- c("Date", "Employment_sector", "Career_stage", "Gender",
                        "Country_of_employment", "Hear_about_workshop",
                        "Workshop", "Have_used_resources_before", 
                        "Will_use_resources_future", "Would_recommend_workshop",
                        "Overall_satisfaction", "Enjoy", "To_improve", 
                        "Long_term", "Browser", "IP", "ID", "LatLon"
                        )
# Data munging
clean_feedback <- feedbackall %>% 
    select(-IP) %>% 
    mutate(Date = lubridate::ymd_hms(as.character(Date), tz = "CET"),
           Employment_sector = str_remove(Employment_sector, 
                                          " \\(University\\)"),
           Country_of_employment = 
               str_replace(Country_of_employment, "Other: ", ""),
           Contact_later = ifelse(is.na(Long_term), "No", "Yes")
           ) %>% 
    select(-Long_term ) %>% 
    separate(Browser, c("Browser", "OS"), sep = " / ") %>% 
    separate(LatLon, c("Latitude", "Longitude"), sep = ", ") %>% 
    mutate(Overall_satisfaction = str_remove(
      feedbackall$Overall_satisfaction, pattern = " *\\(.*")) %>% 
    separate(Browser, c("Browser"), extra = "drop") %>% 
    mutate(Latitude = as.double(Latitude), 
           Longitude = as.double(Longitude))
```

From the above numbers we had a total of `r nrow(clean_feedback)` 
responses to the voluntary and annonymous after-workshop feedback. 

```{r echo=FALSE, message=FALSE}
# customised locations
locations <- clean_feedback %>% 
    select(Workshop) %>% 
    mutate(Locality = str_extract(Workshop, pattern = "\\w+$")) %>% 
    mutate(Locality = str_remove(Locality, pattern = "(.*)_")) %>%   
    select(Locality) 
```

Please notice how much the number of feedback
responses varies among different locations.
```{r echo = FALSE}
knitr::kable(
  table(locations), col.names = c("Location", "Number of responses")
)
```

## Visualising responses
```{r echo=FALSE, message=FALSE}
# Setting a custom ggplot2 theme
theme_clean <- function(){
  theme_bw() +
    theme(axis.text.x = element_text(size = 11),
          axis.text.y = element_text(size = 11),
          axis.title.x = element_text(size = 11, face = "plain"),             
          axis.title.y = element_text(size = 11, face = "plain"),             
          panel.grid.major.x = element_blank(),                                          
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.y = element_blank(),  
          plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = , "cm"),
          plot.title = element_text(size = 14, vjust = 1, hjust = 0.5),
          legend.text = element_text(size = 12, face = "italic"),          
          legend.title = element_blank(),                              
          legend.position = c(0.5, 0.8))
}
```

Comparing the gender distribution of respondents, it is similar to the 
total gender distribution of the total attendees. 
```{r echo = FALSE}
per <- as.data.frame(table(clean_feedback$Gender)/nrow(clean_feedback))
per$Freq <- round(per$Freq * 100, 2)
per$Freq <- paste(per$Freq, c("%", "%", "%"))
knitr::kable(per, col.names = c("Gender", "Percentage of respondents"),
             align = c("l","r"))
```

```{r fig.cap = "Gender distribution", message=FALSE, echo=FALSE, fig.height= 2.5}
library(ggplot2)
ggplot(data = clean_feedback, mapping = aes(x = Gender)) +
    geom_bar(fill = "#1B2944", colour = "darkgrey", width = 0.5) +
    labs(title = "Gender") +
    expand_limits(y = 40) +
    coord_flip() +
    theme_clean()
```

Our aim is to reach all career stages, however it is clear that most attendees 
are PhD candidates.

```{r fig.cap = "Career stages", message=FALSE, echo=FALSE, fig.height=4.5}
car_levels <- c("Undergraduate", "Masters", "PhD candidate",
                  "Postdoctoral researcher", 
                   "Senior academic / Principal investigator", 
                   "Industry scientist", "Other")
clean_feedback %>% 
  mutate(career = parse_factor(Career_stage, car_levels)) %>% 
  ggplot(mapping = aes(x = career)) +
    geom_bar(fill = "#1B2944", colour = "darkgrey", width = 0.5) +
    expand_limits(y = 40) +
    labs(title = "Career stage of respondents",
         x = "Career stage") +
    coord_flip() +
    theme_clean()
```

## Workshop satisfaction

Overall the workshops are well received.

```{r fig.cap = "Workshop feedback", message=FALSE, echo=FALSE, fig.height=3.5}
sat_levels <- c("Satisfactory", "Good",
                  "Very good", "Excellent")
clean_feedback %>% 
  mutate(sat = parse_factor(Overall_satisfaction, sat_levels)) %>% 
  ggplot(mapping = aes(x = sat)) +
    geom_bar(fill = "#1B2944", colour = "darkgrey", width = 0.5) +
    expand_limits(y = 40) +
    labs(title = "Overall satisfaction with the workshop",
         x = "Feedback responses") +
    coord_flip() +
    theme_clean()
```

We also asked if attendees would recommend the workshop, and most people said 
yes.

```{r fig.cap = "Recommendations", message=FALSE, echo=FALSE, fig.height=2.5}
clean_feedback %>% 
  ggplot(mapping = aes(x = Would_recommend_workshop)) +
    geom_bar(fill = "#1B2944", colour = "darkgrey", width = 0.5) +
    labs(title = "Would you recommend the workshop",
         x = "Recommended") +
    coord_flip() +
    theme_clean()
```

## Extra questions

These workshops are all bring-your-own-device. The distribution of operative
systems is interesting. Most attendees use Windows computers.

```{r echo = FALSE}
osatt <- str_extract(clean_feedback$OS, pattern = "\\w+")
knitr::kable(
  table(osatt), col.names = c("Operative System", "Number of laptops"),
  caption = "OS from participants"
)
```

The last two questions were only included from the third workshop and have 
`r sum(!is.na(clean_feedback$Have_used_resources_before))` responses each.

First, we asked if attendees have used the tools or resources before attending the
workshop. 

```{r echo = FALSE}
knitr::kable(
  table(clean_feedback$Have_used_resources_before), 
  col.names = c("How often have you used the tools before", "Number of responses"),
  caption = "Have used the tools or resources before"
)
```

Then, we asked if attendees will use the tools or resources they have learned 
after attending the workshop. 

```{r echo = FALSE}
knitr::kable(
  table(clean_feedback$Will_use_resources_future), 
  col.names = c("Will you use the tools or resources in the future", "Number of responses"),
  caption = "Will use tools or resources after the workshop"
)
```



# Open source

This report was written in Rmarkdown and uses the open-source a default style. 
It has been published using 
Github pages. To see the Rmarkdown source of this document, you may follow this
[link to github](https://github.com/orchid00/ElixirBE/blob/master/reports/2018March.Rmd). 

This report is freely available under [MIT License License](https://github.com/orchid00/ElixirBE/blob/master/LICENSE.md). You may re-use and adapt
the material in any way you wish, without asking permission, **provided you cite the original source**. That is a [link back to this report](https://orchid00.github.io/ElixirBE/reports/2018March.html).

I acknowledge this publication is resulting from support of [Elixir-Belgium](https://www.elixir-belgium.org/) for my role as data science 
and bioinformatics trainer. 

Last update: `r Sys.Date()`


